---
title: "Assignment 4"
format: html
editor: visual
---

```{r}
#| label: setup
#| output: false
#| message: false
#| echo: false
library(eurostat)
library(dineq)
library(readr)
library(tidyverse)
library(stringr)
library(huxtable)
library(vtable)
library(ggplotify)
library(grid)
library(dplyr)
library(gridExtra)
library(ggplot2)
library(broom)
library(lmtest)
library(plm)
```

```{r, show_col_types = FALSE}
#| echo: FALSE
#| output: FALSE
#| label: Datasets and variable mutations
# Loading datasets
# Country code list: "IT", "SE", "BE", "AT", "HR", "RS", "BA", "BG", "DE")
# 
nama_10r_3gdp_linear_csv <- read_csv("nama_10r_3gdp_linear.csv.gz", 
    na = "NA")
demo_r_pjanaggr3_linear_csv <- read_csv("demo_r_pjanaggr3_linear.csv.gz", 
    na = "NA")
roads <- read_csv("roads.gz")
educ <- read_csv("educ.gz")
aream2 <- read_csv("aream2.gz")

# Subset of countries from 2000 to 2020
# Filtering geo = 5 to only get NUTS3 level data
gdp_filtered <- nama_10r_3gdp_linear_csv |> 
  filter(
    nchar(geo) == 5, 
    unit == "MIO_EUR", 
      TIME_PERIOD %in% c(2000:2020),
    str_sub(geo, start = 1L, end = 2L) %in% c("IT", "SE", "BE", "AT", "HR", "RS", "BA", "BG", "DE")
    ) |> 
  select(
    year = TIME_PERIOD, 
    geo, GDP = OBS_VALUE
    ) 

# Filtering population demo for nuts3 countries and years
# Filtering sex = T to only get the total of both genders
# Filtering age = TOTAL.
pop <- demo_r_pjanaggr3_linear_csv |> 
    filter(
      nchar(geo) == 5,
      sex == "T",
      age == "TOTAL",
      TIME_PERIOD %in% c(2000:2020), 
    str_sub(geo, start = 1L, end = 2L) %in% c("IT", "SE", "BE", "AT", "HR", "RS", "BA", "BG", "DE")
    ) |> 
  select(
    year = TIME_PERIOD, 
    geo, Population = OBS_VALUE
    ) 

# Merging the two datasets
merged_data <- merge(gdp_filtered, pop, by = "geo") |> 
  select(
    -year.y,
    Year = year.x,)

# merging the two datasets from eurostat and removing NA and zero values
merged_data <- merge(gdp_filtered, pop, by = c("geo","year"), all = TRUE) |> 
filter(!is.na(Population),
         Population != 0,
         GDP != 0, 
         GDP != "NA"
)
# Generating the GDP per capita variable
merged_data <- merged_data |> 
  mutate(
    GDPpcn3 = GDP / Population
  )
# Adding GDP variable by deviding gdp on population
merged_data <- merged_data |> 
  mutate(GDPpcn3 = (GDP / Population)*1000000)

# Calculating GINI in nuts3 data for nuts2 regions to measure inequity on a nuts2 scale
merged_data <- merged_data |> 
 mutate(
    nuts2 = substr(geo, 1,4),
    nuts0 = substr(geo, 1,2)
 )

# Cleaning the road dataset by removing irrelevant data and NA values
roads <- roads |> 
  select(
    nuts2 = geo, Roadkm = OBS_VALUE, year = TIME_PERIOD
  ) 
roads <- roads |>
  filter(
    !is.na(Roadkm),
    Roadkm != 0
  ) 
# Cleaning education dataset renaming variables
educ <- educ |> 
  filter(sex == "T") |> 
  select(
    nuts2 = geo, age, year = TIME_PERIOD, education = OBS_VALUE, level = isced11
  ) |> 
  filter(nuts2 != "NA", age != "NA", year != "NA", education != "NA")
  
# We are only looking for data for College education (ED5-8) and we devide education by 100 to get the percentages on percentage form.
educ <- educ |> 
  mutate(
    education = education/100
  ) |> 
  filter(level == "ED5-8") 

# Aream2 refers to the dataset containing area data for each nuts3 region measured in squared kilometres
aream2 <- aream2 |>
    mutate(
    year2 = 2010
  ) |> 
   select(
    nuts2 = geo, year = year2, area = OBS_VALUE
  ) 
# We noticed the dataset contained age groups, and two of those groups contained much of the same data: age 25-34 and 30-34. This created issues as some of these observations would be counted twice. We therefore decied to remove the group 30-34.
educ <- educ |> 
  filter(
    !age %in% "Y30-34"
  ) |> 
  mutate(
     nuts2 = substr(nuts2, 1,4),
  ) |> 
  group_by(nuts2) |> 
  mutate(
    sumeduc = sum(education, na.rm = TRUE)
  ) |> 
    slice(1) |> 
    ungroup() |> 
  select(-age)

merged_data <- merged_data |> 
  filter(!is.na(Population),
         Population != 0,
         GDPpcn3 != 0, 
         GDPpcn3 != "NA"
  ) |> 
  mutate(
         GDPpcn2 = mean(GDPpcn3), 
         popn2 = sum(Population),
         GDPn2 = sum(GDP),
         numberofregions = n_distinct(geo)
         ) 
 # select(nuts0, nuts2, geo, year, GDP, Population, GDPpcn3, gini_n2, numberofregions, GDPn2, GDPpcn2, popn2) 


```

# Introduction

Bakgrunn:

Objectives; Outline the main objectives and research questions addressed in Assignments 1, 2, and 3.

**Significance:** Highlight the importance and relevance of the research undertaken.

Snakk om viktigheten til hvorfor se på Gdp og ulikhet, andre ting som kan påvirke som; vei, utdanning, motorvei.

Forklar at noen land er rik og noen er fattig.
For å belyse dette på en god måte bør man bruke land med mange opservasjoner.
Vi ser på Tyskland og Italia pga de har ulik økonomi & industri.

# Literature Review

### 1. Literature review

Regional inequality shows variation in regional GDP per capita, and according to @lessmann2017, regional inequality occurs when one region experiences different assumptions than other regions.
Moreover, when the level of economic development is high, there is significant regional inequality [@lessmann2017].
According to @williamson1965, one crucial difference that leads to regional inequality is the access to natural resources.
Regions rich in natural resources grow faster than regions in the rest of the country, which results in inequality.
Furthermore, openness is another critical element of why regional inequality occurs.
A region with more significant access to the international trade market than other regions with limited access will gain more from international trade if international trade costs decrease [@lessmann2017].

Moreover, openness can also contribute to foreign investors investing in a more attractive region, e.g., access to natural resources.
In the end, this led to an imbalance in regional equality.
Additionally, mobility can provide regional inequality through differences in transport costs from one region to another [@lessmann2017].
The differences in transport costs are caused, e.g. by fuel consumption per kilometre.
 

Furthermore, another significant element creating regional inequality is bureaucracy, policy and stability in a country [@lessmann2017].
Civil war affects the country's balance of regional equality and creates different assumptions.
An example is regions in a country that experience increasing aid.
Receiving assistance from foreign aid organisations creates inequality.
Lastly, @lessmann2017 state that an area where ethnic groups dominate as a minority can result in regional inequity.

Based on the research conducted by @lessmann2017, considering regional inequities is essential since it can influence economic outcomes.
The research shows many examples of high- and low-income countries experiencing unbalanced regional development that results in income inequality [@lessmann2017].

Tidligere arbeid!!
oppsummer hovedlitteraturen til reginal devolopment og inequality

Research gap!
Nytt arbeid

Vi ser på videre utvikling frem til 2020 ettersom denne littraturen slutter i 2012

# Data and Methodology

**Data Sources:** Briefly describe the primary data sources used in the previous assignments, noting any limitations.

### 1. Data acquisition

By using Eurostat, we have downloaded data containing GDP and population for each country in the EU from year 1990 to 2023.
However, we are only interested in the following countries from 2000 to 2020:

-   Austria
-   Belgium
-   Bosnia and Herzegovina
-   Bulgaria
-   Croatia
-   Germany
-   Italy
-   Serbia
-   Sweden

## Description

The first dataset we downloaded from Eurostat was **demo_r_pjanaggr3,** which contains population data from Europe.
In our case, we have selected population data from Sweden, Italy, Belgium, Austria, Croatia, Serbia, Bosnia and Herzegovina, Germany.
According to @eurostat2021, the dataset of demography is a set of statistical numbers each country in Europe sends yearly to Eurostat by 31 December of the following calendar year.
Furthermore, the definition of "population" is, according to @eurostat2021, all persons having their residences in a member state, including all live births and deaths during the year when the statistical numbers are heading to Eurostat.
Moreover, the demographic data at a regional level are separated into different groups named NUTS, which are defined as nomenclature for territorial units for statistics [@eurostat2021].
The purpose of the NUTS function is to subdivide the territory of the European Union into 104 regions at NUTS level 1 and break it down by to 281 at NUTS level 2 and 1348 at NUTS level 3 [@eurostat2021].
In the beginning of this paper, we have focused on the numbers of NUTS 3 level, but the dataset containing population data has been broken down by sex and broad age groups (0-14, 15-64 and 65 or more).

The following dataset we downloaded from Eurostat was the **nama_10r_3gdp dataset**.
Furthermore, this dataset is also broken down to NUTS level 3 which focus on GDP data from the sub-regions in Sweden, Italy, Belgium, Austria, Croatia, Serbia, Bosnia and Herzegovina, Germany.
According to @eurostat2023, GDP (Gross domsetsic product) at market prices is defined as the final result of the production activity of resident producer units [@eurostat2023].

### 2. GDP per capita calculation

In order to visualize the GDP per capita, we have merged the two datasets together and added a new variable that executes this formula in order to generate the GDP per capita for every country.

$y_i = \frac{\text{GDP}_i}{\text{population}_i} * 10^6$

It is worth noting that the GDP data from Eurostat is represented in million euros.
In order to better understand the graphs we have produced, we decided to multiply by one million, so that all numbers in the graph containing the GDP per capita is in full Euros.

GDP or Gross Domestic Product is a measure of value creation in a country.
This is described by Eurostat to be "a measure of economic activity, defined as the value of all goods and services produced minus the value of any goods or services used in their creation" @eurostat2023.
How Eurostat calculates the annual growth rate in GDP "comparisons of the dynamics of economic development both over time and between economies of different sizes" @eurostat2023.
To measure the growth rate of GDP over time "in volume, GDP in current prices is valued at the prices of the previous year, and the volume changes thus calculated are imposed on the level of a reference year; this is called a chain-linked series. Consequently, price movements will not inflate growth" @eurostat2023.

### 3. Descriptive analysis

```{r}
#Legg til alle landene i denne summery tabbelen
#| echo: FALSE
#| tbl-cap: Summary_Austria
# Generates descriptive statistics for variables in plot for AT

merged_data |> 
  filter(str_detect(geo,"AT")) |> 
  select(-1,-year) |> 
  st("GDPpcn3", add.median = TRUE)


```

**Skriv ut i fra denne tabellen som argument for hvilke land vi tar med oss videre!!**

**Methodological Approach:** Summarize the methodologies applied in Assignments 1, 2, and 3, highlighting the rationale behind choosing each method.

De ulike metodene er;

Assigment 1 brukte vi gdp mot gini

Skriv om at vi brukte logaritmic, quadratic + vi må tilføre quic som metode.
Dette gjøres for å se hvilke metode som er "Best fit" og begrunn valg for de ulike metodene.

# Empirical Findings

**Cross-sectional Estimates:** Summarize essential results obtained from the cross-sectional analysis in Assignment 2.

```{r}
#| echo: false
#| cap-location: bottom
#| tbl-cap: "Linear regression model year 2010 for Germany and Italy"
# Estimating linear regression models for Germany and Italy with Gini and GDP per capita.
# By using the formula we observed skewed data and mitigated the issue by log transforming the independant variable. 
lmDE <- lm(gini_n2 ~ log(GDPpcn2), subset(merged_data, year == 2010 & nuts0 == "DE" & gini_n2 > 0))
lmIT <- lm(gini_n2 ~ log(GDPpcn2), subset(merged_data, year == 2010 & nuts0 == "IT" & gini_n2 > 0))

lmSE <- lm(gini_n2 ~ log(GDPpcn2), subset(merged_data, year == 2010 & nuts0 == "SE" & gini_n2 > 0))

huxreg(DE = lmDE, IT = lmIT, SE = lmSE, statistics = c("Number of observations" = "nobs", "r.squared","adj.r.squared", "statistic", "p.value"))
```

In the table above we see the coefficients, standard errors, and significance levels (indicated by asterisks) for a simple linear regression model for the year 2010.
The model has GDP per capita as the independent variable and regional Gini as the dependent variable for both Germany (DE) and Italy (IT).

For Germany (DE), the intercept coefficient is -2.129\*\*, which indicates a high level of statistical significance at the 1% level.
This means that when GDP per capita is zero, the estimated Gini coefficient for Germany is -2.129.
The standard error of the intercept is 0.755, indicating the precision of the estimate.

For Italy (IT), the intercept coefficient is -0.478 and is not significant.
This means that when GDP per capita is zero, the estimated Gini coefficient for Italy is 0.478.
The standard error of the intercept is 0.433.

For Germany (DE), the coefficient for log(GDPpcn2) is 0.222\*\*.
This implies a high level of statistical significance at the 1% level.
In other words, there is strong evidence of a positive linear relationship between the GDP per capita and the Gini coefficient in Germany in 2010.
The standard error associated with the coefficient is 0.073, which indicates the precision of this relationship.

For Italy (IT), the coefficient for log(GDPpcn2) is 0.052.
The absence of asterisks suggests that this coefficient is not statistically significant at commonly used significance levels.
This implies that there is no strong statistical evidence of a linear relationship between the GDP per capita and the Gini coefficient in Italy in 2010.
The standard error of the coefficient is 0.043, providing a measure of the precision of this relationship.

## **Model Diagnostics**

R-squared (r.squared) is a measure of how well the independent variable(s) explain the variation in the dependent variable.
For Germany, the R-squared is 0.229, indicating that approximately 22.9% of the variation in the Gini coefficient can be explained by variations in log(GDPpcn2).
For Italy, the R-squared is 0.090, suggesting that about 9.0% of the variation in the Gini coefficient can be explained by variations in log(GDPpcn2).

Adjusted R-squared (adj.r.squared) is a modified version of R-squared that accounts for the number of independent variables.
It penalizes the inclusion of unnecessary variables.
For both Germany and Italy, the adjusted R-squared is lower than the R-squared, indicating that some variables may not be contributing significantly to the model.

The statistic value (statistic) and p-value (p.value) are typically used to test the overall significance of the regression model.
For Germany, the t-statistic is 9.183, and the associated p-value is 0.005, indicating that the model is statistically significant at the 0.05 significance level.
For Italy, the t-statistic is 1.485, and the associated p-value is 0.242, suggesting that the model is not statistically significant at the 0.05 significance level.

### **Classical assumptions OLS for MLR**

There are underlying assumptions in statistical analysis like ordinary least squares regression (OLS).
OLS yields the best estimates when these conventional presumptions hold true.
OLS regression is based on six traditional assumptions.
Our models are unbiased if the first four conditions are met.
Below is a list of all six presumptions:

1.  Linearity: There must be a linear relationship between the dependent variable (regional GINI) and independent variable (GDP per capita).
    We can read from the two scatterplots below, @fig-scatterplot-RL-DE and @fig-scatterplot-RL-IT, that the relationship for both Germany and Italy seems to be linear, with only some few outliers.

2.  Sample variation: We can also see from @fig-scatterplot-RL-DE and @fig-scatterplot-RL-IT that there is variation in the independent variable (GDP per capita).

3.  Random sampling: There is no random sampling of the population as we are looking at the entire population.

4.  Exogeneity: As inequality is closely related to the development of a country as undeveloped countries typically have higher levels of inequality, there is likely concerns of endogenity.
    This means that there is likely a correlation between the dependent and independent variable.

5.  Homoskedasticity: To fulfill the OLS assumptions, variance in the observations needs to be equal, known as homoskedasticity.
    There are many tests for heteroskedasticity.
    In each case, the null hypothesis is homoskedasticity and the alternative is heteroskedasticity.
    If the variance in the observations are not equal, heteroskedasticity could be present.
    This may result in incorrect standard errors that are computed for the least squares estimators, and could lead to incorrect conclusions.
    In order to check for heteroskedasticity, we utilize the Breusch-Pagan test, **see appendix model X**

6.  Normality: In the context of ordinary least squares (OLS) assumptions, the normality criteria demands that the error term adheres to a normal distribution with an expected value of zero.
    This implies that the discrepancies between observed and predicted values, known as residuals, should have a mean of zero, ensuring a homoskedasticity error term.
    However, it is imperative to note that the failure of the error term to equal zero implies a departure from normality in the population error term within the dataset, thus rendering the normality assumption unfulfilled.

Visualusering av modellen i assignment 2;

```{r}
#| echo: FALSE
#| label: fig-scatterplot-RL-DE
#| cap-location: bottom
#| fig-cap: "Regression model for the year 2010 for Germany with GDP per capita (x) and regional inequality (Gini/y)."
# Visualizing the linear regression of the gini coefficient by GDP per capita for Germany
merged_data_filtered <- merged_data %>%
  filter(str_detect(geo, "DE") & year == 2010 & gini_n2 > 0)

ggplot(merged_data_filtered, aes(x = log(GDPpcn2), y = gini_n2)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x,2), se = F, span = 1) +  
  labs(
    x = "GDP per capita nuts2",
    y = "Gini Index (n2)"
  )
```

In @fig-scatterplot-RL-DE, we plot GDP per capita on nuts 2 level (x) against Gini (y) and illustrate the estimated relationship with a blue line.
The inequality (y-axis) starts off with a slight decrease as the GDP per capita (x-axis) increases.
However, inequality seem to increase.

```{r}
#| echo: FALSE
#| label: fig-scatterplot-RL-IT
#| cap-location: bottom
#| fig-cap: "Regression model for the year 2010 for Italy with GDP per capita (x) and regional inequality (Gini/y)."
# Visualizing the linear regression of the gini coefficient by GDP per capita for Italy
merged_data_filtered <- merged_data %>%
  filter(str_detect(geo, "IT") & year == 2010 & gini_n2 > 0)

ggplot(merged_data_filtered, aes(x = log(GDPpcn2), y = gini_n2)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x,2), se = F, span = 1) +  # Specify the formula
  labs(
    x = "GDP per capita nuts2",
    y = "Gini Index (n2)"
  )
```

In @fig-scatterplot-RL-IT, we plot GDP per capita on nuts 2 level (x) against Gini (y) and illustrate the estimated relationship with a blue line.
Italy has a bigger decrease in inequality compared to Germany.
The inequality (y-axis) starts off with a decrease as the GDP per capita (x-axis) increases.
However, at some point it increases again meaning an increasing GDP per capita result in bigger regional inequality.

Assigment 2; her legges variablene roads, area og educ til.

To further understand the inequality on a regional level it is necessary to look for other variables that can contribute to explaining the Gini coefficient.
Variables that may help explain this is the educational level of the population, the area of the region as well as the amount of highway infrastructure that a region has.
The education data gathered is measured in the percentage of the population that has completed a higher education level (ED5-8) from the age of 20 to 65.
The area data is measured in squared kilometers and the road data is the total amount of highway infrastructure that a region has.

If a region has a big difference in the level of education among subregions, we believe this will increase the regional inequality.
Human capital is one of the most important determinants of differences in reginal development within countries @lessmann_and_seidel2017.
*"Human capital creates positive externalities, which can also spill over reginal boundaries"* @lessmann_and_seidel2017.
Unfortunately, this data is not available for Germany and Italy.
We therefore fall back upon the average quality of human capital within Germany and Italy, as measured by the education level ED5-8 to look if a percentage of higher education increase inequality in a region.

We believe that if a region has a big difference in the level of education among subregions, this will increase the regional inequality based on the assumption that a subregion with higher educated citizens will typically earn more than subregions with lesser educated citizens, thus increasing the difference in GDP per capita on the regional level.
Another assumption that can explain the increased expectancy, is that higher educated citizens tend to have a higher probability of centralizing towards city centers, contrary to less educated citizens.

The centralization assumption is also true for the area of the regions.
If a region is large, the probability for centralization could be bigger, based on factors such as job markets, mobility and distance between cities or areas.
This is the basis of why we chose to add the highway infrastructure variable, as a higher level of infrastructure might decrease the likelihood for centralization as mobility increases.
  

### 2. **Multiple Linear Regression Model**

```{r}
#| echo: FALSE
# Adding the new variables to the main dataset
merged_data <- left_join(merged_data, aream2, by = c("nuts2", "year"))

merged_data <- left_join(merged_data, educ, by = c("nuts2", "year"))

merged_data <- left_join(merged_data, roads, by = c("nuts2", "year"))
```

```{r}
#| echo: FALSE
#| label: Dataset for 2010 variables
# Creating a new dataset that only contains Germany and Italy for the year 2010.
merged2010 <- merged_data |> 
  filter(year == 2010) |> 
  select(nuts2, year, everything()) |> 
  select(nuts2, year, GDPpcn2, gini_n2, education, area, Roadkm) |> 
  filter(str_detect(nuts2, "IT|DE"), nchar(nuts2) == 4) |> 
  filter(
    Roadkm != "NA",
    Roadkm != 0
  ) |> 
  mutate(
    nuts0 = substr(nuts2, 1,2)
  ) |> 
    mutate(
    roadratio = Roadkm/area
  ) |> 
  select(
    nuts0, nuts2, everything()
  )
```

```{r}
#| echo: FALSE
#| tbl-cap: Multiple regression models

# In lm1 and lm2 we look at how the GDP per capita of each nuts 2 region as well as education may explain the gini coefficient. We have used. a log transformation on the GDP because both gini and education is measured in percentages, and without doing so the GDP data is significantly skewed

lm1 <- lm(gini_n2 ~ log(GDPpcn2) + education, subset(merged2010, year == 2010 & nuts0 == "DE" & gini_n2 > 0))
lm2 <- lm(gini_n2 ~ log(GDPpcn2) + education, subset(merged2010, year == 2010 & nuts0 == "IT" & gini_n2 > 0))

# lmarea looks at the above, but adds the area variable. This is also log transformed as we are interested in the changes in percentage

lmarea <- lm(gini_n2 ~ log(GDPpcn2) + education + log(area), subset(merged2010, year == 2010 & nuts0 == "DE" & gini_n2 > 0))

lmarea2 <- lm(gini_n2 ~ log(GDPpcn2) + education + log(area), subset(merged2010, year == 2010 & nuts0 == "IT" & gini_n2 > 0))

# lm3 and lm4 also includes roads (amount of highway in kilometres in a region). These models takes all explanatory variables in to account.

lm3 <- lm(gini_n2 ~ log(GDPpcn2) + education + log(area) + log(Roadkm), subset(merged2010, year == 2010 & nuts0 == "DE" & gini_n2 > 0))

lm4 <- lm(gini_n2 ~ log(GDPpcn2) + education + log(area) + log(Roadkm), subset(merged2010, year == 2010 & nuts0 == "IT" & gini_n2 > 0))

huxreg("Model 1 DE" = lm1, "Model 1 IT" = lm2, "Model 2 DE" = lmarea, "Model 2 IT" = lmarea2, "Model 3 DE" = lm3, "Model 3 IT" = lm4, statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
```

The provided multiple regression models aim to explain regional inequality, with three different models.

-   Model 1; Explains how regional development (GDP per capita) as the independent variable and regional inequality (gini_n2) as the dependent variable is affected by education when it comes to inequality for the country's DE (Germany), and IT (Italy).
-   Model 2; Explains how regional development (GDP per capita) as the independent variable and regional inequality (gini_n2) as the dependent variable is affected by education and area when it comes to inequality for the country's DE (Germany), and IT (Italy).
-   Model 3; Explains how regional development (GDP per capita) as the independent variable and regional inequality (gini_n2) as the dependent variable is affected by education + area + Road(km) when it comes to inequality for the country's DE (Germany), and IT (Italy).

### 3. **Model Interpretation**

In Model 1, the R-squared values are 0.222 for Germany and 0.122 for Italy.
In Model 2, they increase to 0.309 for Germany and 0.338 for Italy.
In Model 3, they remain relatively consistent at 0.313 for Germany and 0.340 for Italy.

The adjusted R-squared values follow a similar pattern, indicating that Model 2 has the highest explanatory power for both Germany and Italy.
However, the increase in R-squared between Model 1 and Model 2 suggests that Model 2 provides a better fit compared to Model 1.

The coefficients associated with each independent variable (log(GDPpcn2), education, log(area), and log(Roadkm)) indicate the direction and strength of their influence on regional inequality.
For log(GDPpcn2), positive coefficients indicate that an increase in GDP per capita is associated with higher regional inequality.
In Model 2 and Model 3, log(GDPpcn2) is statistically significant at the 0.01 significance level for both Germany and Italy.
For education, negative coefficients suggest that higher education levels are associated with lower regional inequality.
However, in all models, education is not statistically significant.
For log(area), positive coefficients imply that larger regions tend to have higher regional inequality.
In Model 2 and Model 3, log(area) is statistically significant at the 0.05 significance level for both Germany and Italy.
For log(Roadkm), the coefficients are generally small, and in Model 2 and Model 3, they are not statistically significant.

The F-statistic value and p-value indicate that Model 2 is statistically significant at the 0.01 significance level for both Germany and Italy, while Model 1 is statistically significant for Germany but not for Italy.
Model 3 is statistically significant at the 0.05 significance level for Germany but not for Italy.

In summary, Model 2 appears to have the best overall fit, as indicated by the highest R-squared and adjusted R-squared values for both Germany and Italy.
The primary contributor to explaining variations in regional inequality appears to be log(GDPpcn2), as it is consistently statistically significant at the 0.01 significance level in Model 2 and Model 3.
The impact of other independent variables, such as education, log(area), and log(Roadkm, is less clear, as they are generally not statistically significant in the models. The specific implications of these findings would depend on the context of your analysis and the objectives of your research. It's important to consider the practical significance of these results and conduct further analysis if needed.

**Alternative Functional Forms and Panel Estimates:** Present the key insights derived from exploring alternative functional forms and panel estimates in Assignment 3.

# Discussion

**Key Insights:** Synthesize and discuss the central insights emerging from the empirical findings.

**Policy Implications:** Deliberate on the potential implications of your results for policy-making and practice in the realm of regional development and inequality.

# Limitations and Future Research

**Research Limitations:** Acknowledge any constraints or limitations encountered during your analysis.

# Conclusion

**Summary:** Provide a succinct summary of the main findings and their significance.

**Final Reflection:** Briefly reflect on the research process and the contributions of your study to the field of regional development and inequality.

# References

```{r}

```
